---
title: "IMDB Feature Film Analysis"
author: "T2 Deep Learners"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
options(scientific=T, digits = 3) 

library(tidyr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(ggthemes)
library(ezids)

```

# Data cleaning and Merging

```{r data_join}

movies <- read.csv('movies.csv', header=TRUE)
drop <- c('budget','usa_gross_income','worlwide_gross_income','metascore','reviews_from_users',
          'reviews_from_critics','production_company','description','writer')
movies <- movies[,!names(movies) %in% drop]
movies <- movies[!(movies$country == "") & !(movies$language == "") & !(movies$director == "") & !(movies$actors == ""),]

ratings <- read.csv('ratings.csv', header=TRUE)
drop <- c('allgenders_0age_avg_vote', 'allgenders_0age_votes', 'males_0age_avg_vote', 'males_0age_votes',
          'females_0age_avg_vote', 'females_0age_votes', 'us_voters_rating', 'us_voters_votes', 
          'non_us_voters_rating', 'non_us_voters_votes')
ratings <- ratings[,!names(ratings) %in% drop]
ratings <- na.omit(ratings)

movie_ratings <- merge(movies, ratings, by="imdb_title_id" )

```

# EDA

### Movie Rating and QQ plot

```{r movie_rating_&_qq_plot}
movie_genre_1 <- movie_ratings[,c("genre", "avg_vote")]

data_mg <- aggregate(movie_genre_1$avg_vote, list(movie_genre_1$genre), FUN=mean, sort = FALSE)
names(data_mg)[names(data_mg) == "Group.1"] <- "genre"
#count_mg <- movie_genre_1%>%count(genre)
names(data_mg)[names(data_mg) == "x"] <- "rating"
names(data_mg)[names(data_mg) =="n"] <- "count"
data_mg$genre <- as.factor(data_mg$genre)

# summary of dataset
summary(data_mg$rating)

# plot histogram plot
hist(data_mg$rating, main = "Histogram of the movie rating", xlab="Movie rating", col="#66C2A5", breaks = 15)

# plot qqplot
qqnorm(data_mg$rating, main = "Q-Q plot for movie rating")
qqline(data_mg$rating)


#remove outliers
outliers <- unique(boxplot(data_mg$rating, plot=FALSE)$out)
data_mg <- data_mg[-which(data_mg$rating %in% outliers),]

# plot ggplot for rating

data_mg <- subset(data_mg, rating>7.5)
ggplot(data=data_mg, aes(x=reorder(genre, rating), y=rating)) + 
  geom_bar(stat = "identity", position="dodge",alpha=.8, fill = "#66C2A5") + 
    scale_fill_fermenter(palette = "Set2") + 
      xlab("Movie genre") + 
        ylab("Movie Rating")+ 
          labs(title='The rating of various movie genres (rating >= 7.5)') + 
            coord_flip() + 
              theme(text = element_text(size=10), legend.position="right", plot.title = element_text(size=15))

```

### Total movie votes and QQ plot

```{r movie votes & qq plot}
movie_genre_v <- movie_ratings[,c("genre", "votes")]
data_v <- aggregate(movie_genre_v$votes, list(movie_genre_v$genre), FUN=sum, sort = FALSE)
names(data_v)[names(data_v) == "Group.1"] <- "genre"
names(data_v)[names(data_v) == "x"] <- "movie_votes"
names(data_v)[names(data_v) =="n"] <- "count"
data_v$genre <- as.factor(data_v$genre)

# summary of dataset
summary(data_v$movie_votes)

# plot histogram plot
hist(data_v$movie_votes, main = "Histogram of the movie votes", xlab="Movie votes", col="#FC8D62", breaks = 15)

# plot qq plot
qqnorm(data_v$movie_votes, main = "Q-Q plot for movie votes")
qqline(data_v$movie_votes)

#remove outliers
outliers <- unique(boxplot(data_v$movie_votes, plot=FALSE)$out)
data_v <- data_v[-which(data_v$movie_votes %in% outliers),]


# ggplot for movie votes
data_v <- subset(data_v, movie_votes > 200000)
ggplot(data=data_v, aes(x=reorder(genre, movie_votes), y=movie_votes)) + 
  geom_bar(stat = "identity", alpha=.7,fill = "#FC8D62") + 
    scale_fill_fermenter(palette = "Set2") + 
      xlab("Movie genre") + ylab("Movie Votes")+ 
        labs(title='The sum of movie votes in movie genres (votes >= 200k)') + 
          coord_flip()+
            theme(axis.text = element_text(size=8), legend.position="right", plot.title = element_text(size=15))

```
### Year-wise movie distribution

```{r EDA_year_movie_count, warning=F, fig.height=4, fig.width=6}

# average vote distribution
movies<-movies[!movies$year=='TV Movie 2019' & !movies$year=='2020',]

# A brief overview of the dataset: the minimum vote is 1.000,while the maximum vote is 9.900.
# the mean of the vote is 5.902, the median is 6.100.


# movies year-numbers distribution
years_numbers <- movies[,c('imdb_title_id', 'year')]
years_num <- years_numbers %>% count(year,sort = TRUE)
colnames(years_num)<-c('Year','Movie_numbers')
years_num$Year <- as.numeric(years_num$Year)
df1 <-arrange(years_num, -Movie_numbers)
head(df1)
tail(df1)


ggplot(years_num,aes(x=Year,y=Movie_numbers)) + 
  geom_bar(stat="identity", fill='#66C2A5', alpha=.8, width=1) + 
    labs(title=('Count of movies each year')) + 
      scale_fill_fermenter(palette = "Set2") + 
        ylab('Count') +
          xlab('Year') +
            coord_flip() +
              scale_x_continuous(breaks=seq(1900, 2020, 10))

```

### Average votes per year

```{r EDA_year_avg_votes, warning=F, fig.height=4, fig.width=6}

# movies year-rating distribution
years_rating <- aggregate(movies$avg_vote, by = list(movies$year),FUN=mean)
colnames(years_rating) <- c('Year','Rating')
years_rating$Year <- as.numeric(years_rating$Year)
df2 <-arrange(years_rating, -Rating)
head(df2)
tail(df2)


ggplot(data=years_rating,aes(x=Year,y=Rating)) +
  geom_bar(stat="identity", fill='#FC8D62', alpha=.8, width=1) +
    scale_fill_fermenter(palette = "Set2") + 
      labs(title=('Average Rating every Year')) + 
        ylab('Count') +
          xlab('Year') +
            coord_flip() +
              scale_x_continuous(breaks=seq(1900, 2020, 10))

```


### Number of votes per year

```{r EDA_year_votes, fig.height=4, fig.width=6}

# movies vote distribution
years_vote <- aggregate(movies$votes, by = list(director = movies$year),FUN=sum)
colnames(years_vote) <- c('Year','Votes')
years_vote$Year <- as.numeric(years_vote$Year)
df3 <-arrange(years_vote, -Votes)
head(df3)
tail(df3)

ggplot(data=years_vote, aes(x=Year, y=Votes)) +
  geom_bar(stat="identity", fill='#66C2A5', alpha=.8, width=1) +
    scale_fill_fermenter(palette = "Set2") + 
      labs(title=('Total Votes every Year')) + 
        ylab('Votes') +
          xlab('Year') +
            coord_flip() +
              scale_x_continuous(breaks=seq(1900, 2020, 10))

```

### Duration of movies per year

```{r EDA_year_duration, fig.height=4, fig.width=6}

# movies duration-year distribution
years_duration <- aggregate(movies$duration, by = list(director = movies$year),FUN=mean)
colnames(years_duration) <- c('Year','Duration')
years_duration$Year <- as.numeric(years_duration$Year)
df4 <-arrange(years_duration, -Duration)
head(df4)
tail(df4)


ggplot(data=years_duration, aes(x=Year, y=Duration)) +
  geom_bar(stat="identity", fill='#FC8D62', alpha=.8, width=1) +
    scale_fill_fermenter(palette = "Set2") + 
      labs(title=('Years v/s Movie Duration')) + 
        ylab('Duration') +
          xlab('Year') +
            coord_flip() +
              scale_x_continuous(breaks=seq(1900, 2020, 10))

```


# Difference in movie Ratings across genres

### Dataset Creation

```{r Q1_dataset_creation, warning=F}

genre_df1 <- movie_ratings[,c("genre", "avg_vote")]

# remove outliers
outliers <- unique(boxplot(genre_df1$avg_vote, plot=FALSE)$out)
genre_df1 <- genre_df1[-which(genre_df1$avg_vote %in% outliers),]

genre_df2 <- genre_df1 %>%
  separate(genre, c("c1",'c2','c3'),sep =c(', '))

df1 <- genre_df2[,c("c1", "avg_vote")] %>% drop_na()
df2 <- genre_df2[,c("c2", "avg_vote")] %>% drop_na()
df3 <- genre_df2[,c("c3", "avg_vote")] %>% drop_na()

names(df1)[1] <- "genre"
names(df2)[1] <- "genre"
names(df3)[1] <- "genre"

genre_df <- rbind(df1, df2, df3)

```

### Data Visualization

```{r Q1_visualization, warning=F, fig.height=5, fig.width=7}

theme_set(theme_bw())

genre_grouped <- genre_df %>%
  group_by(genre) %>%
  filter(n() >= 30) %>%
    summarize(
      mean_vote = mean(avg_vote),
      n=n()
    )

genre_grouped <- genre_grouped[order(genre_grouped$mean_vote), ]
genre_grouped$genre <- factor(genre_grouped$genre, levels = genre_grouped$genre)

ggplot(genre_grouped, aes(fill=mean_vote, y=mean_vote, x=genre)) +
    geom_bar(stat="identity", alpha=.8, width=0.8) +
      scale_fill_fermenter(palette = "Set2") +
        coord_flip() +
            labs(title="Ordered Bar Chart",
                subtitle="Mean avg_vote across film genres", 
                caption="genre: avg_vote")

```

### Hypothesis Testing and P-value visualization

```{r Q1_hypo_test, warning=F}

genre_df_filtered <- genre_df %>%
  group_by(genre) %>%
    filter(n() >= 30) %>%
      ungroup()

pw_ttest <- pairwise.t.test(
  x=genre_df_filtered$avg_vote, 
  g=genre_df_filtered$genre,
  p.adjust.method="bonferroni"
)

pw_ttest_pvals <- data.frame(pw_ttest$p.value)
pw_ttest_pvals$genre1 <- rownames(pw_ttest_pvals)
pw_ttest_pvals <- melt(pw_ttest_pvals, id.vars="genre1", variable.name="genre2", value.name="p_value") %>% drop_na()

ggplot(pw_ttest_pvals, aes(genre1, genre2, fill= p_value)) + 
  geom_tile() +
    scale_fill_gradient(low="darkblue", high = "lightblue") +
      theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
        labs(title="Heat Map",
            subtitle="P-values for pairwise T-tests on movie genres", 
            caption="genre2: genre1")

paste0("Since most p-values are far lesser than our alpha-value of 0.01, we can say that the genre definitely affects the avg_vote.")
```


# Choice of Director affect movie ratings

### EDA

```{r Q2, fig.height=5, fig.width=7}

movie_director <- movie_ratings[,c("director", "avg_vote")]

# remove outliers
outliers <- unique(boxplot(movie_director$avg_vote, plot=FALSE)$out)
movie_director <- movie_director[-which(movie_director$avg_vote %in% outliers),]

director_group <- aggregate(movie_director$avg_vote, by = list(director = movie_director$director),FUN=mean)
director_num <- movie_director %>% count(director,sort = TRUE)
director_vote <- merge(director_group,director_num,by='director')
director_vote <- director_vote[order(-director_vote$n),]
colnames(director_vote) <- c('director','rating','movie_numbers') 

director_vote_50 <- subset(director_vote,movie_numbers>49)
director_vote_50$rating <- as.numeric(director_vote_50$rating)

ggplot(director_vote_50, aes(fill=rating, y=rating, x=reorder(director,rating))) +
    geom_bar(stat="identity", alpha=.8, width=0.8) +
      scale_fill_fermenter(palette = "Set2") +
        coord_flip() +
            labs(title="Ordered Bar Chart",
                subtitle="Mean avg_vote for various Directors",
                caption="director: avg_vote")

```

### Hypothesis Testing and P-value visualization

```{r Q2_hypo_test, warning=F}

movie_director_filtered <- movie_director %>%
  group_by(director) %>%
    filter(n() >= 30) %>%
      ungroup()

anovaRes = aov(avg_vote ~ director, data=movie_director_filtered)
anovaRes 
summary(anovaRes)

tukeyDirectorAoV <- TukeyHSD(anovaRes)
tukeyVar <- data.frame(tukeyDirectorAoV$director)
tukeyVarSig <- subset(tukeyVar, p.adj <= 0.01)
head(tukeyVarSig)

paste0("We see a list of director pairs where the movie ratings distribution for their movies are significantly different.")
paste0("Hence, we can say that the choice of director is important in selecting a movie.")

```

# Movie genre viewing statistic differences with age

### Dataset Creation

```{r Q3_dataset_creation, warning=F}

genre_age_df1 <- movie_ratings[,c("genre", "allgenders_18age_votes", "allgenders_30age_votes", "allgenders_45age_votes")]

genre_age_df2 <- genre_age_df1 %>%
    mutate(
      prop_18 = allgenders_18age_votes / (allgenders_18age_votes + allgenders_30age_votes + allgenders_45age_votes),
      prop_30 = allgenders_30age_votes / (allgenders_18age_votes + allgenders_30age_votes + allgenders_45age_votes),
      prop_45 = allgenders_45age_votes / (allgenders_18age_votes + allgenders_30age_votes + allgenders_45age_votes)
    )

names(genre_age_df2)[2:4] <- c("votes_18", "votes_30", "votes_45")

genre_age_df3 <- genre_age_df2 %>%
  separate(genre, c("c1",'c2','c3'),sep =c(', '))

df1 <- genre_age_df3[,c("c1", "votes_18", "votes_30", "votes_45", "prop_18", "prop_30", "prop_45")] %>% drop_na()
df2 <- genre_age_df3[,c("c2", "votes_18", "votes_30", "votes_45", "prop_18", "prop_30", "prop_45")] %>% drop_na()
df3 <- genre_age_df3[,c("c3", "votes_18", "votes_30", "votes_45", "prop_18", "prop_30", "prop_45")] %>% drop_na()

names(df1)[1] <- "genre"
names(df2)[1] <- "genre"
names(df3)[1] <- "genre"

genre_age_df <- rbind(df1, df2, df3)

```


### Data Visualization

```{r Q3_visualization, warning=F}
ggplot(genre_age_df, aes(x=prop_18, group=genre, fill=genre)) +
  geom_density(adjust=1.5, alpha=.4) +
      ggtitle("Distribution of Age 18 viewing proportions for various genres")

ggplot(genre_age_df, aes(x=prop_30, group=genre, fill=genre)) +
  geom_density(adjust=1.5, alpha=.4) +
      ggtitle("Distribution of Age 30 viewing proportions for various genres")

ggplot(genre_age_df, aes(x=prop_45, group=genre, fill=genre)) +
  geom_density(adjust=1.5, alpha=.4) +
      ggtitle("Distribution of Age 45 viewing proportions for various genres")

genre_age_grouped <- genre_age_df %>%
  group_by(genre) %>%
    summarize(
      ages_18 = mean(prop_18),
      ages_30 = mean(prop_30),
      ages_45 = mean(prop_45)
    )

genre_age_melt <- melt(genre_age_grouped, id.vars = "genre", variable.name = "proportions")

ggplot(genre_age_melt, aes(fill=proportions, y=value, x=genre)) + 
    geom_bar(position="stack", stat="identity") +
        scale_fill_brewer(palette = "Set2") +
            coord_flip() + 
                ggtitle("Mean of proportions of viewer ages across film genres")

paste0("We see how proportion differences across various genres differ for different ages.")

```

### Hypothesis Testing

```{r Q3_hypo_test, warning=F}

genre_age_df_filtered <- genre_age_df %>%
  group_by(genre) %>%
    filter(n() >= 30) %>%
      ungroup()

res.man <- manova(cbind(votes_18, votes_30, votes_45) ~ genre, data = genre_age_df_filtered)
summary(res.man)
paste0("A very low p-value tells us that genre differences are significant across various ages.")

summary.aov(res.man)
paste0("A low p-value for each age category tells us that all the ages have differing viewing statistics across different genres.")

```

# Affect of duration of movie on avg_vote

### EDA

```{r Q4_EDA}

# movies duration-rating distribution
duration <-outlierKD2(movies, movies$duration)
outliers <- boxplot(movies$duration,plot=FALSE)$out
movies1 <- movies[-which(movies$duration %in% outliers),]
summary(movies1$duration)

duration_rating <- movies1[,c('duration','avg_vote')]
colnames(duration_rating)<-c("duration","rating")
df5 <-arrange(duration_rating, -rating)


duration_grouped <- duration_rating %>%
  group_by(duration) %>%
    summarize(
      mean_vote = mean(rating),
      n=n()
    )

ggplot(duration_grouped, aes(fill=mean_vote, y=mean_vote, x=duration)) +
    geom_bar(stat="identity", alpha=.8, width=0.8) +
      scale_fill_fermenter(palette = "Set2") +
        coord_flip() +
            labs(title="Ordered Bar Chart",
                subtitle="Mean rating for different film durations",
                caption="duration: Rating")

```

### Correlation Analysis

```{r Q4_corr_analysis}

#correlation 
cor.test(movies$avg_vote, as.numeric(movies$duration))
ggplot(movies,aes(x=duration, y=avg_vote)) + 
  scale_fill_fermenter(palette = "Set2") +
    ylab('Rating') +
      geom_point() +
        geom_smooth(method='lm') 

paste0("There is a weak correlation between movie duration and Rating.")

```

# Effect of movie language on Average Vote

### Dataset Creation

```{r Q5_dataset_creation, warning=F}

language_df1 <- movie_ratings[,c("language", "avg_vote")]

# remove outliers
outliers <- unique(boxplot(language_df1$avg_vote, plot=FALSE)$out)
language_df1 <- language_df1[-which(language_df1$avg_vote %in% outliers),]

language_df2 <- language_df1 %>%
  separate(language, c("c1",'c2','c3',"c4",'c5','c6'),sep =c(', '))

df1 <- language_df2[,c("c1", "avg_vote")] %>% drop_na()
df2 <- language_df2[,c("c2", "avg_vote")] %>% drop_na()
df3 <- language_df2[,c("c3", "avg_vote")] %>% drop_na()
df4 <- language_df2[,c("c4", "avg_vote")] %>% drop_na()
df5 <- language_df2[,c("c5", "avg_vote")] %>% drop_na()
df6 <- language_df2[,c("c6", "avg_vote")] %>% drop_na()

names(df1)[1] <- "language"
names(df2)[1] <- "language"
names(df3)[1] <- "language"
names(df4)[1] <- "language"
names(df5)[1] <- "language"
names(df6)[1] <- "language"

language_df <- rbind(df1, df2, df3, df4, df5, df6)

```

### Data Visualization

```{r Q5_visualization, warning=F, fig.height=5, fig.width=10}

theme_set(theme_bw())

language_grouped <- language_df %>%
  group_by(language) %>%
    filter(n() >= 200) %>%
      summarize(
        mean_vote = mean(avg_vote),
        n=n()
      )

ggplot(language_grouped, aes(y=mean_vote, x=reorder(language,mean_vote))) +
    geom_point(width=.8, stat="identity") +
      scale_fill_fermenter(palette = "Set2") +
          xlab("language") +
          theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
            labs(title="Ordered Scatter Plot",
                subtitle="Mean vote across film languages (>= 200 votes)", 
                caption="language: mean_vote")

```

### Hypothesis Testing and P-value visualization

```{r Q5_hypo_test, warning=F}

language_df_filtered <- language_df %>%
  group_by(language) %>%
    filter(n() >= 30) %>%
      ungroup()

anovaRes = aov(avg_vote ~ language, data=language_df_filtered)
anovaRes 
summary(anovaRes)
names(anovaRes)

tukeyLanguageAoV <- TukeyHSD(anovaRes)
tukeyVar <- data.frame(tukeyLanguageAoV$language)
tukeyVarSig <- subset(tukeyVar, p.adj <= 0.01)
head(tukeyVarSig)

```

# Voter demographics affecting average vote and votes

### Average Vote v/s Gender

```{r Q6_movie_rating_vs_gender, warning=F}

# rating
mean_all <- rowMeans(subset(movie_ratings[, c(28, 30, 32)]))
mean_male <- rowMeans(subset(movie_ratings[, c(36, 38, 40)]))
mean_female <- rowMeans(subset(movie_ratings[, c(44, 46, 48)]))

movie_gender = data.frame(movie_ratings[,c("genre")], mean_all, mean_male, mean_female)
movie_gender <- aggregate(list(movie_gender$mean_all, movie_gender$mean_male, movie_gender$mean_female), by = list(movie_gender$`movie_ratings...c..genre...`), mean)
colnames(movie_gender) <- c("genre","all gender","male", "female")

# t-test for movie rating
t.test(movie_gender$male, movie_gender$female)


# barplot for movie ratings for both genders
colnames(movie_gender) <- c("Genre","All gender","Male", "Female")
barplot(colMeans(movie_gender[,3:4]), col = c("#FC8D62", "#8DA0CB"), main="The average of movie rating in female and male")

# melt data frame into long
colnames(movie_gender) <- c("genre","all gender","male", "female")
movie_gender <- melt(movie_gender)

# boxplot: movie ratings vs. genders
boxplot(value~variable, data=movie_gender, main="Movie rating vs. genders", xlab="Gender", ylab="Movie rating", col=c("#66C2A5", "#FC8D62", "#8DA0CB")) 

```

### Total Votes v/s Gender

```{r movie_vote_vs_gender}

# vote dataset with genders
vote_all <- subset(movie_ratings[, c(35,43)])
vote = data.frame(movie_ratings$genre, vote_all$males_allages_votes, vote_all$females_allages_votes)

vote <- aggregate(list(vote$vote_all.males_allages_votes, vote$vote_all.females_allages_votes), by = list(vote$movie_ratings.genre), sum)
colnames(vote) <- c("Genre","Male", "Female")


# t-test for movie votes
t.test(vote$Male, vote$Female)

paste0("Difference in mean number of voters is significant between male and female genders.")


# plot histogram plot for male
hist(vote$Male, main = "Histogram of the movie votes for male", xlab="Movie rating", col="#FC8D62", breaks = 15)

# plot qqplot for male
qqnorm(vote$Male, main = "Q-Q plot for movie votes for males")
qqline(vote$Male)


# plot histogram plot for female
hist(vote$Female, main = "Histogram of the movie votes for female", xlab="Movie rating", col="#8DA0CB", breaks = 15)

# plot qqplot for female
qqnorm(vote$Female, main = "Q-Q plot for movie votes for females")
qqline(vote$Female)

# bar plot: movie votes vs. genders
barplot(colSums(vote[,2:3]), col = c("#FC8D62", "#8DA0CB"), main="The total votes for females and males")

# melt data frame into long
colnames(vote) <- c("genre","male", "female")
vote <- melt(vote)

# boxplot: genders vs. movie votes
boxplot(value~variable, data=vote, main="Movie votes vs. genders", xlab="Gender",ylab="Movie rating",col=c("#FC8D62", "#8DA0CB"), border=c("#FC8D62", "#8DA0CB"))

```

### Dataset Creation

```{r Q6_dataset_creation, warning=F}

age_df <- movie_ratings[,c("allgenders_18age_avg_vote", "allgenders_30age_avg_vote", "allgenders_45age_avg_vote", "avg_vote")]
names(age_df)[1:3] <- c("avg_vote_18", "avg_vote_30", "avg_vote_45")

# remove outliers
outliers_18 <- unique(boxplot(age_df$avg_vote_18, plot=FALSE)$out)
age_df <- age_df[-which(age_df$avg_vote_18 %in% outliers_18),]

outliers_30 <- unique(boxplot(age_df$avg_vote_30, plot=FALSE)$out)
age_df <- age_df[-which(age_df$avg_vote_30 %in% outliers_30),]

outliers_45 <- unique(boxplot(age_df$avg_vote_45, plot=FALSE)$out)
age_df <- age_df[-which(age_df$avg_vote_45 %in% outliers_45),]

gender_df <- movie_ratings[,c("males_allages_avg_vote", "females_allages_avg_vote", "avg_vote")]
names(gender_df)[1:2] <- c("avg_vote_male", "avg_vote_female")

# remove outliers
outliers_male <- unique(boxplot(gender_df$avg_vote_male, plot=FALSE)$out)
gender_df <- gender_df[-which(gender_df$avg_vote_male %in% outliers_male),]

outliers_female <- unique(boxplot(gender_df$avg_vote_female, plot=FALSE)$out)
gender_df <- gender_df[-which(gender_df$avg_vote_female %in% outliers_female),]

```



### Data Visualization

```{r Q6_visualization, warning=F, fig.width=6, fig.height=4}

age_df_cor <- cor(age_df)
age_df_melt <- melt(age_df_cor)

ggplot(age_df_melt, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile(aes(label = value)) +
    geom_text(aes(label = round(value,2))) +
      theme(axis.title.x=element_blank(), axis.title.y=element_blank()) +
        scale_fill_fermenter(palette = "Blues") +
          labs(title="Heat Map",
              subtitle="Age Correlation Matrix",
              caption="age: avg_vote")

gender_df_cor <- cor(gender_df)
gender_df_melt <- melt(gender_df_cor)

ggplot(gender_df_melt, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
    geom_text(aes(label = round(value,2))) +
      theme(axis.title.x=element_blank(), axis.title.y=element_blank()) +
        scale_fill_gradient(low="darkblue", high = "lightblue") +
          labs(title="Heat Map",
              subtitle="Gender Correlation Matrix",
              caption="gender: avg_vote")

```

### Hypothesis Testing 

```{r Q6_hypo_test, warning=F}

ttest_age_18_30 <- t.test(age_df$avg_vote_18, age_df$avg_vote_30, conf.level = 0.99)
ttest_age_18_30

paste0("Difference in average votes is significant between ages 18 and 30.")

ttest_age_30_45 <- t.test(age_df$avg_vote_30, age_df$avg_vote_45, conf.level = 0.99)
ttest_age_30_45

paste0("Difference in average votes is significant between ages 30 and 45.")

ttest_age_45_18 <- t.test(age_df$avg_vote_45, age_df$avg_vote_18, conf.level = 0.99)
ttest_age_45_18

paste0("Difference in average votes is significant between ages 18 and 45.")

ttest_gender <- t.test(gender_df$avg_vote_male, gender_df$avg_vote_female, conf.level = 0.99)
ttest_gender

paste0("Difference in average votes is significant between male and female genders.")

```